---
title: "p8105_hw2_oew2106"
author: "Olivia Wagner"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1 ##
```{r problem 1, Data set 1}

# Read in Data Set #
Trash_Wheel_Data = readxl::read_excel('./Problem 1/Trash-Wheel-Collection-Totals-8-6-19.xlsx', col_names = TRUE, na = '', range = 'A2:N406')

# alter the names of the columns to snake format#
Trash_Wheel_Data = janitor::clean_names(Trash_Wheel_Data)


# filter out the results without dumpster information, and round the number of sports balls found #
Trash_Wheel_Data = Trash_Wheel_Data %>% filter(!is.na(dumpster)) %>% mutate(sports_balls = as.integer(sports_balls)) 


# Summary of the Data #
size = nrow(Trash_Wheel_Data)

avg_trash_by_month = Trash_Wheel_Data %>% group_by(month, year) %>% summarise(n = n(), total_weight = sum(weight_tons))

avg_trash = mean(pull(avg_trash_by_month, total_weight))

median_sports_balls = median(pull(Trash_Wheel_Data, sports_balls))

```


```{r problem 1, Data set 2}
# Extract 2017/2018 Data #
precipitation_2018 = readxl::read_excel('./Problem 1/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = 5, col_names = TRUE, na = '', range = 'A2:B14') 

precipitation_2017 = readxl::read_excel('./Problem 1/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = 6, col_names = TRUE, na = '', range = 'A2:B14')

# Alter Data Sets #
precipitation_2018 = precipitation_2018 %>% mutate(year = '2018') %>% filter(!is.na(Total))

precipitation_2017 = precipitation_2017 %>% mutate(year = '2017') %>% filter(!is.na(Total))

# Combine Data Sets #
precipitation_data = rbind(precipitation_2017, precipitation_2018)

# Alter the Month from numerical to character representation #
precipitation_data = precipitation_data %>% mutate(month = month.name[Month], total = Total) %>% select(-Total, -Month) 


# Data Summary #

yearly_precip_data = precipitation_data %>% group_by(year) %>% summarize(sum_precipitation = sum(total))
rainfall_total_2017 = yearly_precip_data %>% filter(year == 2017) %>% pull(sum_precipitation)
rainfall_total_2018 = yearly_precip_data %>% filter(year == 2018) %>% pull(sum_precipitation)
```

  The data in Problem 1 is collected by the Trash Wheel in Baltimore, Maryland. The Wheel collects waste from the harbor, after which the city collects and incinerates the trash as a method of generating electricity for the city. The amount of rainfall is a determinant of the volume of trash collected. There are `size ` observations within the Trash Wheel Data, spanning every month since May of 2014. The average number of tons brought in by the Trash Wheel is `avg_trash `. The total precipitation in the year 2017 was `rainfall_total_2017 `, while the total yearly precipitation for 2018 was `rainfall_total_2018 `. The data also contains information on the different types of trash picked up by the Trash Wheel. Objects such as cigarette butts and sports balls are counted when clearing trash from the water. During 2017 the median number of sports balls Mr. Trash Wheel picked up was `median_sports_balls `.      


## Problem 2 ##
```{r problem 2, Data set 1}

# Read in Data Set #
pols_data = read_csv('./Problem 2/fivethirtyeight_datasets/pols-month.csv')
snp_data = read_csv('./Problem 2/fivethirtyeight_datasets/snp.csv')
unemployment_data = read_csv('./Problem 2/fivethirtyeight_datasets/unemployment.csv')

# Clean pols_data #
pols_data = pols_data %>% separate(mon, into = c('year', 'month', 'day'), sep = '-') %>% mutate(month = as.integer(month)) %>% mutate(month = month.abb[month]) %>% mutate(president = ifelse(prez_gop == 1, 'gop', 'dem')) %>% select(-prez_gop, -prez_dem, -day) 

# Clean snp_data #

snp_data = snp_data %>% separate(date, into = c('month', 'day', 'year'), sep = '/') %>% mutate(year = as.numeric(year), month = as.numeric(month), day = as.numeric(day)) %>% select(year, month, close) %>%  arrange(year, month) %>% mutate(month  = month.abb[month]) 

# Clean unemployment_data #

unemployment_data = unemployment_data %>% gather(Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, key = 'month', value = 'rate') %>% group_by(Year, month) %>% arrange(Year) %>% rename(year = Year, unemp_rate = rate)

# Merge the Data Frames/ preserve order #

pol_snp_data = merge(pols_data, snp_data, by = c('year', 'month'))
total_538_data = merge(pol_snp_data, unemployment_data, by = c('year', 'month')) 

```

  
  The data table is comprised of data taken from three sources, the first a categorization of political offices by their party membership, the second a list of S&P market values at the close of each month, and finally the percentage of unemployment by month. All of these data sets contain information from January 1950 or prior. The resulting data table is 786 rows long 11 columns long , and is organized by year and month columns. 


## Problem 3 ##
```{r problem 3, Data set 1}

# Read in Data Set #

name_data = read_csv('./Problem 3/Popular_Baby_Names.csv') 
name_data = janitor::clean_names(name_data)

print(name_data)
# Alter Ethnicity category to streamline formatting #
name_data = name_data %>% mutate(ethnicity = replace(ethnicity, ethnicity == 'ASIAN AND PACI', 'ASIAN AND PACIFIC ISLANDER')) %>% mutate(ethnicity = replace(ethnicity, ethnicity == 'BLACK NON HISP', 'BLACK NON HISPANIC')) %>% mutate(ethnicity = replace(ethnicity, ethnicity == 'WHITE NON HISP', 'WHITE NON HISPANIC'))

# Alter the Name cateogry to ensure the same formatting #

name_data = name_data %>% mutate(childs_first_name = tolower(childs_first_name)) %>% print()
```
